A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-02 kobakian-cook.tex
Potentially use NSW for thyroid example?

In many countries, residents are increasingly gathering to live in urban
areas around major cities \citep{ACTUC}. The creation of approximately
equal population areas results in dramatically different square meterage
of the geographic units. The geographic units are usually
administrative, such as states or electorates. The units are filled with
colour to represent the value of the statistic for each unit \citep{EI}.
This can lead to unequal emphasis on the statistical value of each of
the geographic units. This allows choropleth map displays to
misrepresent the spatial distributions of human related statistics due
to area-size bias \citep{BCM}. Figure \ref{fig:melanoma-geo} is a
choropleth map that uses colour to display the estimated Standardised
Incidence Ratios (SIRs) of melanoma cancer for males for each of the
Statistical Areas at Level 2 (SA2) used by the Australian Bureau of
Statistics \citeyearpar{abs2011}. The Australian choropleth map display
draws attention to the expanse of light blue areas across the rural
communities in all states. The SA2s around Brisbane stand out as more
orange and red. To create this choropleth map the SA2 polygons for 2011
from the Australian Bureau of Statistics.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_melanoma_p.pdf}
\caption{\label{fig:melanoma-geo}A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}

@@

There are several established alternative visualisation methods. Tile
maps, Rectangular cartograms \citep{ORC} and Dorling cartograms
\citep{ACTUC} all use one simple shape to represent each geographic
unit. They all minimise the emphasis on the size or shape of the
geographic areas. These alternative map displays focus on the
relationship between neighbours, attempting to preserve connections, and
disregard the unique shapes of the administrative boundaries. Figure
\ref{fig:qld_grid} shows a collection of alternative map displays, this
includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a
Dorling cartogram and d) a hexagon tile map. Each Statistical Area at
Level 2 (SA2) \citep{abs2011} was designed to represent a community.
These administrative areas are often used to aggregate census data and
used to understand demographics of the Australian population.

The familiar choropleth map display is often used to allow the user to
orient themselves. However, the emphasis on the land mass diminishes the
features of the distribution in densely populated communities due to the
small size on the display \citep{ACTUC}. The unique shapes of boundaries
can be helpful for orienting users but may not contribute to their
understanding of the spatial disease distribution as many of the
communities are not visible in a choropleth display \citep{TVSSS}. This
presents an opportunity for explicit map transformations to improve
communication of distributions \citep{CBATCC}. In figure
\ref{fig:qld_grid}a) the choropleth map is presented with the
alternative maps to allow for comparisons to be made.

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/qld_grid.pdf}
\caption{\label{fig:qld_grid}A collection of alternative maps of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, ranging from much lower than average (blue) to much higher than average (red)}
\end{figure}

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The impact of a disease over
connected communities can be seen by allowing the boundaries to maintain
connection in the transformed display. The contiguous cartogram
displayed in figure \ref{fig:qld_grid}b) draws attention to smaller
geographic units when they are rescaled according to the population
\citep{DMAHP}. These new shapes can now be coloured to represent a
second variable. This display can create twisted and unfamiliar shapes
from the geographic units as the algorithms must satisfy the topology
conditions, especially when there are communities located geographically
far from their neighbours \citep{TVSSS}. This shown in the northern SA2s
of Queensland with an average (yellow) estimated SIR of Melanoma for
males.

The non-contiguous cartogram in figure \ref{fig:qld_grid}c) also uses
the population to rescale the geographic units. Unlike the contiguous
cartogram, the SA2 areas maintain their geographic shape, but they may
not retain the connection with their neighbours. The population of the
SA2 areas is used to scale the geographic units in the non-contiguous
cartogram \citep{NAC}. The amount of white space can be meaningful in
non-contiguous cartograms \citep{ECGC}, in this example the disparity
between large rural areas and small urban areas means that the city of
Brisbane, shown in the inset map looks reasonable, but these units are
not visible in the context of the Queensland state map. Depending on the
difference between the population and geographic land size, the amount
of white space can also prevent meaningful understanding of the
distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:qld_grid}d) shows the Queensland SA2
areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Queensland that were not apparent in figure
\ref{fig:qld_grid} a), b) or c).

@@

\begin{figure}
\includegraphics[width=14cm]{figs/flowchart.pdf}
\caption{\label{fig:sugarbag_flow}A flow diagram detailing the necessary steps to create a hexagon tile map.}
\end{figure}

@@

The following code creates the hexagon tile map for all the Statistical
Areas at Level 2 in Queensland. These steps can be executed by the main
function, \texttt{create\_hexmap}, or can be implemented separately for
more flexibility.

@@

\begin{Schunk}
\begin{Sinput}
## Load data
qld_sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "Queensland")

## Create centroids set
centroids <- create_centroids(qld_sa2, "sa2_name_2011")

## Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.12,
  ## same size used in create_grid
  hex_filter = 10,
  use_neighbours = qld_sa2,
  focal_points = capital_cities %>%
    filter(points == "Brisbane"),
  width = 35,
  verbose = FALSE
)

fort_qld <- qld_sa2 %>%
  fortify_sfc()

## Make a plot
library(ggplot2)
qld_hexmap <- ggplot() +
  geom_polygon(aes(
      x = long,
      y = lat,
      group = interaction(sa2_name_2011, polygon)
    ),
    fill = "white",
    colour = "lightgrey",
    data = fort_qld
  ) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = hex_id,
    fill = rownumber
  ),
  data = fort_hex) +
  scale_fill_distiller("", palette = "PRGn") +
  coord_equal() + theme_void()
qld_hexmap
\end{Sinput}
\end{Schunk}

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_gghexmap.pdf}
\caption{\label{fig:melanoma-hex}A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}

@@

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2011} of
Queensland in 2011 is provided with the \texttt{sugarbag} package as
\texttt{tas\_sa2}. A single column of the data set is used to identify
the unique areas. In this case, the unique SA2 names for each SA2 have
been used.

The longitude and latitude centre of the capital cities of Australia are
used as focal points to allocate each geographic area around the closest
capital city. Brisbane will be the common focal point, as this example
uses only the areas in the state of Queensland.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/1centroids.pdf}
\caption{\label{fig:centroids_plot}The geographic shapes of the Statistical Areas of Queensland at Level 2. The points show the locations of the centroids of the SA2 areas.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/2grid.pdf}
\caption{\label{fig:grid2}Grid points to create a hexagon tile map.}
\end{figure}

@@

The grid points are kept only if they fall between the rolling average
of the minimum and maximum centroid values after accounting for the
buffer distance, for each row and column of the grid. The sparsely
populated South-West region of National Park has much fewer points
available compared to the South-East region containing the city of
Brisbane.

\begin{figure}[h]
\centering
\includegraphics[width=10cm]{figs/3grid.pdf}
\caption{\label{fig:filter-grid}The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step.}
\end{figure}

@@

The distance between each centroid in the set, and each of the focal
points provided is calculated. The name of the closest focal point, and
the distance and angle from focal point to polygon centroid is joined to
polygon data set. To minimise time taken for this step only one option
is provided, Queensland's capital city Brisbane. The order for
allocation is determined by the distance between the polygon centroid
and it's closest focal point. The points are arranged from the centroid
closest to the focal point(s), to the furthest.

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/4grid.pdf}
\caption{\label{fig:buffers}The remaining available hexagon locations after filtering for the grid points within a square distance, then circular around the centroid.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/5allocate.pdf}
\caption{\label{fig:angles}The remaining available hexagon locations after filtering for grid points within the angle from the focal point to the centroid.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/6allocate.pdf}
\caption{\label{fig:buffs}A complete hexagon tile map of Queensland.}
\end{figure}

A complete hexagon tile map of Queensland is created by applying the
algorithm steps to each centroid. The hexagon tile map visualisation is
used below to visualise the Australian Cancer Atlas data \citep{TACA}.
Two views of the same data are produced by filling according to the Lung
Cancer Standardised Incidence Rates (SIRs) downloaded from the
Australian Cancer Atlas site. This small example in Figure \ref{fig:sir}
shows the group of blue areas in the Brisbane CBD more prominently in
the hexagon tile map (b). The small red areas visible in the choropleth
map (a) along the north coast are much larger in the hexagon tile maps.
The hexagon tile map shows less yellow, this no longer overwhelms the
map space with the information regarding the rural areas.

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/7SIR.pdf}
\caption{\label{fig:sir}The Australian Cancer Atlas data displayed on a choropleth and hexagon tile map. The colour of each Statistical Area of Australian at Level 2 coloured by SIR. A choropleth map (a) of SIR is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented.}
\end{figure}

@@

This algorithm has been applied to the complete set of all Australian
Statistical Areas at Level 2. This display highlights the density of
Australian capital cities, as it draws attention to the many communities
in Sydney, Melbourne and Brisbane. This display also highlights the
disparity in the burden of Melanoma cancer for males in the communities
of these three cities. There are several collections of red hexagons in
Brisbane that draw attention and they represent the communities with
much higher rates of diagnosis than the Australian average. The
communities south of Brisbane show a gradient of colour as the values
progress from higher than average, to much lower than average for the
communities closer to the Sydney CBD. This pattern progresses into the
communities with lower than average rates in Melbourne and Tasmania.

Compared to the choropleth map display, the low rates in the rural
Australian communities are no longer dominating the display. The much
higher than average rates in Sydney draw more attention in the hexagon
tile map display unlike the red and orange areas in the city of Brisbane
that draw attention in both displays.

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_gghexmap.pdf}
\caption{\label{fig:melanoma-geo}A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, they range from much lower than average (blue) to much higher than average (red).}
\end{figure}

@@

The \texttt{gganimate} \citep{gganimate} package can be used to make an
animation. It requires connecting the polygons for each area in two
displays, which can be done using the \texttt{sf\_id} variable, such as
the SA2 name. The animation\footnote{This animation can be viewed at:
  \url{https://sugarbagjss.netlify.com/}} connecting these two displays
will highlight the rapid growth of the inner-city areas that emphasises
the density of the communities. The hexagons that move the furthest will
move rapidly in the animation. The rapid decreases of the large rural
areas also show how much greater the landmass of Statistical Areas can
be.

@@

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \texttt{sugarbag} \citep{sugarbag} package written for \texttt{R}
\citep{R}. The \texttt{sugarbag} package creates tessellated hexagon
tile maps. The Australian application preserves the spatial
relationships, emphasising capital cities. The hexagon tile map is a
visualisation solution that highlights spatial distributions.

These hexagons equally represent each area. However, the tessellation
does not allow the size of the hexagons to represent another variable,
similar to the choropleth maps. The algorithm is heavily dependent on
the focal points used, as this determines the order of allocation. It
works on the assumption that viewers can use directional relationships
to identify their neighbourhoods but this can be aided by the animation.
With careful consideration of the choropleth map, the small geographic
inner city areas may have been noticed by viewers, but the hexagon tile
map display emphasises them. The communities in northern Queensland and
the Northern territory do not draw attention because of their size as in
the choropleth, but their colour is still noticeably below average when
contrasted with the hexagons further south.

The Dorling cartogram was the most influential in the design of the
hexagon tile map. The use of a hexagon provides a simple shape that
shifts the focus of the visualisation from the shape of SA2s to the
distribution effects of a disease on communities. not impeded by the
irregular and unusual shapes created by a contiguous cartogram and the
immense white space created on the non-contiguous cartogram. Hexagons
can emphasise the relationship between neighbouring communities by
allowing their sides to touch. The tessellation crated by adjoining
hexagons for densely populated areas is helpful as it shows that most
communities are not islands onto themselves, but are intrinsically
connected to those around them. As seen in the Dorling cartogram in
figure \ref{fig:qld_grid}c), the tessellation that allows boundary
connections was not imposed for rural communities, where centroid of the
geographic unit was located far from others, a feature that allows
distant but dense populations of the coastal towns of Queensland to be
recognised.

>>>>

In many countries, residents are increasingly gathering to live in urban
areas around major cities \citep{ACTUC}. The creation of approximately
equal population areas results in dramatically different square meterage
of the administrative geographic units, such as states or electorates.
The units are filled with colour to represent the value of the statistic
for each unit \citep{EI}. This can lead to unequal emphasis on the
statistical value of each of the geographic units. This allows
choropleth map displays to misrepresent the spatial distributions of
human related statistics due to area-size bias \citep{BCM}. Figure
\ref{fig:melanoma-geo} is a choropleth map that uses colour to display
the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer
for males for each of the Statistical Areas at Level 2 (SA2) used by the
Australian Bureau of Statistics \citeyearpar{abs2011}. The Australian
choropleth map display draws attention to the expanse of light blue
areas across the rural communities in all states. The SA2s around Hobart
stand out as more orange and red. To create this choropleth map the SA2
polygons for 2011 from the Australian Bureau of Statistics.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_melanoma_p.png}
\caption{\label{fig:melanoma-geo}A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}

@@

There are several established alternative visualisation methods. Tile
maps, Rectangular cartograms \citep{ORC} and Dorling cartograms
\citep{ACTUC} all use one simple shape to represent each geographic
unit. They all minimise the emphasis on the size or shape of the
geographic areas. These alternative map displays focus on the
relationship between neighbours, attempting to preserve connections, and
disregard the unique shapes of the administrative boundaries. Figure
\ref{fig:nsw_grid} shows a collection of alternative map displays, this
includes a) a contiguous cartogram, b) a non-contiguous cartogram, c) a
Dorling cartogram and d) a hexagon tile map. Each Statistical Area at
Level 2 (SA2) \citep{abs2011} was designed to represent a community.
These administrative areas are often used to aggregate census data and
used to understand demographics of the Australian population.

The familiar choropleth map display is often used to allow the user to
orient themselves. However, the emphasis on the land mass diminishes the
features of the distribution in densely populated communities due to the
small size on the display \citep{ACTUC}. The unique shapes of boundaries
can be helpful for orienting users but may not contribute to their
understanding of the spatial disease distribution as many of the
communities are not visible in a choropleth display \citep{TVSSS}. This
presents an opportunity for explicit map transformations to improve
communication of distributions \citep{CBATCC}. In figure
\ref{fig:nsw_grid}a) the choropleth map is presented with the
alternative maps to allow for comparisons to be made.

\begin{Schunk}
\begin{Sinput}
states <- absmapsdata::state2011 %>% 
  rmapshaper::ms_simplify(keep = 0.25, keep_shapes = TRUE) %>% 
  filter(!state_name_2011 == "Other Territories")

sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "New South Wales") %>% 
  filter(!st_is_empty(geometry))

# sa2 <- sa2 %>% rmapshaper::ms_simplify(keep = 0.05, keep_shapes = TRUE)

SIR <- read_csv("data/SIR Downloadable Data.csv") %>% 
  filter(SA2_name %in% sa2$sa2_name_2011) %>% 
  dplyr::select(Cancer_name, SA2_name, Sex_name, p50) %>% 
  filter(Cancer_name == "Thyroid", Sex_name == "Females")
ERP <- read_csv("data/ERP.csv") %>%
  filter(REGIONTYPE == "SA2", Time == 2011, Region %in% SIR$SA2_name) %>% 
  dplyr::select(Region, Value)

# Alternative maps
# Join with sa2 sf object
sa2thyroid_ERP <- SIR %>% 
  left_join(sa2, ., by = c("sa2_name_2011" = "SA2_name")) %>%
  left_join(., ERP %>% 
              dplyr::select(Region, 
              Population = Value), by = c("sa2_name_2011"= "Region")) %>% 
  filter(!st_is_empty(geometry))

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf()

aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  #geom_sf(fill = "lightgrey", data = states) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity()
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
###############################################################################

sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112)
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

###############################################################################
# Contiguous Cartograms

cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()

save(cont, file = "data/cont.rda")
load("data/cont.rda")

aus_ggcont <- ggplot(cont) + 
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
aus_ggcont

ggsave(filename = "figures/aus_ggcont.png", plot = aus_ggcont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
###############################################################################
# Non - Contiguous Cartograms

ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1/10,
                         weight = "Population") %>% st_as_sf()

save(ncont, file = "data/ncont.rda")
load("data/ncont.rda")

aus_ggncont <- ggplot(ncont) + 
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
aus_ggncont

ggsave(filename = "figures/aus_ggncont.png", plot = aus_ggncont,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
###############################################################################
# Non - Contiguous Dorling Cartograms

dorl <- sa2lung_ERP %>%
  mutate(pop = (Population/max(Population))*10) %>% 
  cartogram_dorling(., k = 0.01, weight = "pop", m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity()+
  invthm + guides(fill = FALSE)
aus_ggdorl

ggsave(filename = "figures/aus_ggdorl.png", plot = aus_ggdorl,
       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
## Load data
tas_sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "New South Wales")

## Create centroids set
centroids <- create_centroids(tas_sa2, "sa2_name_2011")

## Allocate polygon centroids to hexagon grid points
hex_allocated <- allocate(
  centroids = centroids,
  hex_grid = grid,
  sf_id = "sa2_name_2011",
  ## same column used in create_centroids
  hex_size = 0.12,
  ## same size used in create_grid
  hex_filter = 10,
  use_neighbours = tas_sa2,
  focal_points = capital_cities %>%
    filter(points == "Hobart"),
  width = 35,
  verbose = FALSE
)

fort_tas <- tas_sa2 %>%
  fortify_sfc()

## Make a plot
library(ggplot2)
tas_hexmap <- ggplot() +
  geom_polygon(aes(
      x = long,
      y = lat,
      group = interaction(sa2_name_2011, polygon)
    ),
    fill = "white",
    colour = "lightgrey",
    data = fort_tas
  ) +
  geom_polygon(aes(
    x = long,
    y = lat,
    group = hex_id,
    fill = rownumber
  ),
  data = fort_hex) +
  scale_fill_distiller("", palette = "PRGn") +
  coord_equal() + theme_void()
tas_hexmap
\end{Sinput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
###############################################################################

aus_grid <- gridExtra::grid.arrange(aus_ggchoro, aus_ggcont, aus_ggncont, aus_ggdorl, nrow = 2)

aus_grid
ggsave(filename = "figures/aus_grid.png", plot = aus_grid,
  device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)
\end{Sinput}
\end{Schunk}

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/nsw_grid.png}
\caption{\label{fig:nsw_grid}A collection of alternative maps of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, ranging from much lower than average (blue) to much higher than average (red)}
\end{figure}

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The impact of a disease over
connected communities can be seen by allowing the boundaries to maintain
connection in the transformed display. The contiguous cartogram
displayed in figure \ref{fig:nsw_grid}b) draws attention to smaller
geographic units when they are rescaled according to the population
\citep{DMAHP}. These new shapes can now be coloured to represent a
second variable. This display can create twisted and unfamiliar shapes
from the geographic units as the algorithms must satisfy the topology
conditions, especially when there are communities located geographically
far from their neighbours \citep{TVSSS}. This shown in the northern SA2s
of New South Wales with an average (yellow) estimated SIR of Melanoma
for males.

The non-contiguous cartogram in figure \ref{fig:nsw_grid}c) also uses
the population to rescale the geographic units. Unlike the contiguous
cartogram, the SA2 areas maintain their geographic shape, but they may
not retain the connection with their neighbours. The population of the
SA2 areas is used to scale the geographic units in the non-contiguous
cartogram \citep{NAC}. The amount of white space can be meaningful in
non-contiguous cartograms \citep{ECGC}, in this example the disparity
between large rural areas and small urban areas means that the city of
Hobart, shown in the inset map looks reasonable, but these units are not
visible in the context of the New South Wales state map. Depending on
the difference between the population and geographic land size, the
amount of white space can also prevent meaningful understanding of the
distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:nsw_grid}d) shows the New South
Wales SA2 areas as an individual circle located as close as possible to
the geographic centroid location. This map draws attention to the
collection of coastal cities in New South Wales that were not apparent
in figure \ref{fig:nsw_grid} a), b) or c).

@@

\begin{figure}
\includegraphics[width=14cm]{figs/flowchart.png}
\caption{\label{fig:sugarbag_flow}A flow diagram detailing the necessary steps to create a hexagon tile map.}
\end{figure}

@@

The following code creates the hexagon tile map for all the Statistical
Areas at Level 2 in New South Wales. These steps can be executed by the
main function, \texttt{create\_hexmap}, or can be implemented separately
for more flexibility.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_gghexmap.png}
\caption{\label{fig:melanoma-hex}A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}

@@

The polygon set of Statistical Areas at Level 2 (SA2) \citep{abs2011} of
New South Wales in 2011 is provided with the \texttt{sugarbag} package
as \texttt{tas\_sa2}. A single column of the data set is used to
identify the unique areas. In this case, the unique SA2 names for each
SA2 have been used.

The longitude and latitude centre of the capital cities of Australia are
used as focal points to allocate each geographic area around the closest
capital city. Hobart will be the common focal point, as this example
uses only the areas in the state of New South Wales.

@@

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/1centroids.png}
\caption{\label{fig:centroids_plot}The geographic shapes of the Statistical Areas of New South Wales at Level 2. The points show the locations of the centroids of the SA2 areas.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/2grid.png}
\caption{\label{fig:grid2}Grid points to create a hexagon tile map.}
\end{figure}

@@

The grid points are kept only if they fall between the rolling average
of the minimum and maximum centroid values after accounting for the
buffer distance, for each row and column of the grid. The sparsely
populated South-West region of National Park has much fewer points
available compared to the South-East region containing the city of
Hobart.

\begin{figure}[h]
\centering
\includegraphics[width=10cm]{figs/3grid.png}
\caption{\label{fig:filter-grid}The remaining hexagon locations from all the possible points in the initial grid after the buffer is applied. The blue dots show the grid points left to choose from after the buffer step.}
\end{figure}

@@

The distance between each centroid in the set, and each of the focal
points provided is calculated. The name of the closest focal point, and
the distance and angle from focal point to polygon centroid is joined to
polygon data set. To minimise time taken for this step only one option
is provided, New South Wales's capital city Hobart. The order for
allocation is determined by the distance between the polygon centroid
and it's closest focal point. The points are arranged from the centroid
closest to the focal point(s), to the furthest.

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/4grid.png}
\caption{\label{fig:buffers}The remaining available hexagon locations after filtering for the grid points within a square distance, then circular around the centroid.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/5allocate.png}
\caption{\label{fig:angles}The remaining available hexagon locations after filtering for grid points within the angle from the focal point to the centroid.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=16cm]{figs/6allocate.png}
\caption{\label{fig:buffs}A complete hexagon tile map of New South Wales.}
\end{figure}

A complete hexagon tile map of New South Wales is created by applying
the algorithm steps to each centroid. The hexagon tile map visualisation
is used below to visualise the Australian Cancer Atlas data
\citep{TACA}. Two views of the same data are produced by filling
according to the thyroid Cancer Standardised Incidence Rates (SIRs)
downloaded from the Australian Cancer Atlas site. This small example in
Figure \ref{fig:sir} shows the group of blue areas in the Hobart CBD
more prominently in the hexagon tile map (b). The small red areas
visible in the choropleth map (a) along the north coast are much larger
in the hexagon tile maps. The hexagon tile map shows less yellow, this
no longer overwhelms the map space with the information regarding the
rural areas.

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/7SIR.png}
\caption{\label{fig:sir}The Australian Cancer Atlas data displayed on a choropleth and hexagon tile map. The colour of each Statistical Area of Australian at Level 2 coloured by SIR. A choropleth map (a) of SIR is paired with a hexagon tile map (b) to contrast the colours that are made obvious when every SA2 is equally represented.}
\end{figure}

@@

This algorithm has been applied to the complete set of all Australian
Statistical Areas at Level 2. This display highlights the density of
Australian capital cities, as it draws attention to the many communities
in Sydney, Melbourne and Hobart. This display also highlights the
disparity in the burden of Melanoma cancer for males in the communities
of these three cities. There are several collections of red hexagons in
Hobart that draw attention and they represent the communities with much
higher rates of diagnosis than the Australian average. The communities
south of Hobart show a gradient of colour as the values progress from
higher than average, to much lower than average for the communities
closer to the Sydney CBD. This pattern progresses into the communities
with lower than average rates in Melbourne and New South Wales.

Compared to the choropleth map display, the low rates in the rural
Australian communities are no longer dominating the display. The much
higher than average rates in Sydney draw more attention in the hexagon
tile map display unlike the red and orange areas in the city of Hobart
that draw attention in both displays.

\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_gghexmap.png}
\caption{\label{fig:melanoma-geo}A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, they range from much lower than average (blue) to much higher than average (red).}
\end{figure}

@@

The \texttt{gganimate} \citep{gganimate} package can be used to make an
animation. It requires connecting the polygons for each area in two
displays, which can be done using the \texttt{sf\_id} variable, such as
the SA2 name. The animation connecting these two displays will highlight
the rapid growth of the inner-city areas that emphasises the density of
the communities. The hexagons that move the furthest will move rapidly
in the animation. The rapid decreases of the large rural areas also show
how much greater the landmass of Statistical Areas can be.

@@

Alternative maps highlight the value of the statistics across the
geographic units. Alternative mapping methods allow increased
understanding of the spatial distribution of a variable across the
population, by fairly representing each administrative area. This
acknowledges that the amount of residents can be different but
recognises that each population area is equally important. The solution
to this visualisation problem has equally sized areas, with
neighbourhood boundary connections. This map algorithm is implemented in
the \texttt{sugarbag} \citep{sugarbag} package written for \texttt{R}.
The \texttt{sugarbag} package creates tessellated hexagon tile maps. The
Australian application preserves the spatial relationships, emphasising
capital cities. The hexagon tile map is a visualisation solution that
highlights spatial distributions.

These hexagons equally represent each area. However, the tessellation
does not allow the size of the hexagons to represent another variable,
similar to the choropleth maps. The algorithm is heavily dependent on
the focal points used, as this determines the order of allocation. It
works on the assumption that viewers can use directional relationships
to identify their neighbourhoods but this can be aided by the animation.
With careful consideration of the choropleth map, the small geographic
inner city areas may have been noticed by viewers, but the hexagon tile
map display emphasises them. The communities in northern New South Wales
and the Northern territory do not draw attention because of their size
as in the choropleth, but their colour is still noticeably below average
when contrasted with the hexagons further south.

The Dorling cartogram was the most influential in the design of the
hexagon tile map. The use of a hexagon provides a simple shape that
shifts the focus of the visualisation from the shape of SA2s to the
distribution effects of a disease on communities. not impeded by the
irregular and unusual shapes created by a contiguous cartogram and the
immense white space created on the non-contiguous cartogram. Hexagons
can emphasise the relationship between neighbouring communities by
allowing their sides to touch. The tessellation crated by adjoining
hexagons for densely populated areas is helpful as it shows that most
communities are not islands onto themselves, but are intrinsically
connected to those around them. As seen in the Dorling cartogram in
figure \ref{fig:nsw_grid}c), the tessellation that allows boundary
connections was not imposed for rural communities, where centroid of the
geographic unit was located far from others, a feature that allows
distant but dense populations of the coastal towns of New South Wales to
be recognised.
